---
title: "12 In-class Exercise (Review)"
author: "Heng Kuan Xin"
date: 2024-11-11
date-modified: "last-modified"

toc: true
toc-expand: true
number-sections: true

execute:
  eval: false
  echo: true
  freeze: true
  output: true
  warning: false
  error: false
---

# Notes on Configuring App Inputs

-   Note that while you can allow users to have the freedom to choose the inputs, such as how much % of the data set to sample and use for regression analysis, to use for test or train data, you should try out the different possible inputs and observe how long the app will take to run for the user.

-   Remember to always set seed whenever you are doing random selection or any random process, so that your results are reproducable, for yourself and the user.

# In-class Exercise

```{r}
pacman::p_load(sf, tidyverse)
```

## Multicollinearity Check

```{r}
#| eval: false

mdata_nogeo <- mdata %>% st_drop_geometry() # no need geometry

ggstatsplot::ggcorrmat(mdata_nogeo[,2:17])
```

## Model callibration

Find the suitable bandwidth first, before applying it to the gwr regression model.

```{r}
#| eval: false
gwr_bw_train_ad <- bw.gwr(resale_price ~ floor_area_sqm +
                            storey_order + remaining_lease_mths +
                            PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                            PROX_MRT + PROX_PARK + PROX_MALL +
                            PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                            WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + WITHIN_1KM_PRISCH,
                          data=train_data,
                          approach="CV",
                          kernel="gaussian",
                          adaptive=TRUE,
                          longlat=FALSE)
```

## Predictive Modelling with RF Method

```{r}
#| eval: false
# extract coordinates from your samples
coords <- st_coordinates(HDB_sample)
coords_train <- st_coordinates(train_data)
coords_test <- st_coordinates(test_data)
```

```{r}
#| eval: false

# Random Forest Model
rf <- ranger(resale_price ~ floor_area_sqm +
                            storey_order + remaining_lease_mths +
                            PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                            PROX_MRT + PROX_PARK + PROX_MALL +
                            PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                            WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + WITHIN_1KM_PRISCH,
                          data=train_data_nogeom)
```

```{r}
#| eval: false
rf_pred <- predicT(rf,
                   data = test_data_nogeom)

rf_pred_df <- as.data.frame(rf_pred$predictions) %>%
  rename(rf_pred = "rf_pred$predictions")
```

```{r}

```
