---
title: "11 'Calibrating Hedonic Pricing Model for Private Highrise Property with GWR Method'"
author: "Heng Kuan Xin"
date: 2024-10-21
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  freeze: true
  output: true
  warning: false
  error: false

toc: true
toc-depth: 4
---

# Overview

(Copied from Exercise):

**Geographically weighted regression (GWR)** is a spatial statistical technique that takes non-stationary variables into
consideration (e.g., climate; demographic factors; physical environment characteristics) and models the local relationships between these independent variables and an outcome of interest (also known as dependent variable). In this hands-on exercise, you will learn how to build [hedonic pricing](https://www.investopedia.com/terms/h/hedonicpricing.asp) models by using GWR methods. The dependent variable is the resale prices of condominium in 2015. The independent variables are divided into either structural and locational.

## Data

Two data sets will be used in this model building exercise, they are:

-   URA Master Plan subzone boundary in shapefile format (i.e. *MP14_SUBZONE_WEB_PL*)

-   condo_resale_2015 in csv format (i.e. *condo_resale_2015.csv*)

## Packages

Before we get started, it is important for us to install the necessary R packages into R and launch these R packages into R environment.

The R packages needed for this exercise are as follows:

-   R package for building OLS and performing diagnostics tests: [**olsrr**](https://olsrr.rsquaredacademy.com/)

-   R package for calibrating geographical weighted family of models: [**GWmodel**](https://cran.r-project.org/web/packages/GWmodel/index.html)

-   R package for multivariate data visualisation and analysis: [**corrplot**](https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html)

-   Spatial data handling: **sf**

-   Attribute data handling: **tidyverse**, which contains the following packages that we want: **readr**, **ggplot2** and **dplyr**

-   Choropleth mapping: **tmap**

```{r}
# Importing the packages
pacman::p_load(olsrr, corrplot, ggpubr, sf, spdep, GWmodel, tmap, tidyverse, gtsummary)
```

```{r}
mpsz.sf <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL") %>% 
  st_transform(crs=3414)

condo_resale <- read_csv("data/aspatial/Condo_resale_2015.csv")

condo_resale.sf <- condo_resale %>% 
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"),
            crs=4326) %>%
  st_transform(crs=3414)
```

## Exploratory Data Analysis

```{r}
st_bbox(mpsz.sf) #view the bounding box
```

```{r}
# head(condo_resale.sf)
# head(mpsz.sf)
```

```{r}
ggplot(data=condo_resale.sf, aes(x=`SELLING_PRICE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")
```

```{r}
condo_resale.sf <- condo_resale.sf %>%
  mutate(`LOG_SELLING_PRICE` = log(SELLING_PRICE))

ggplot(data=condo_resale.sf, aes(x=`LOG_SELLING_PRICE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")
```

```{r}
AREA_SQM <- ggplot(data=condo_resale.sf, aes(x= `AREA_SQM`)) + 
  geom_histogram(bins=20, color="black", fill="light blue")

AGE <- ggplot(data=condo_resale.sf, aes(x= `AGE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_CBD <- ggplot(data=condo_resale.sf, aes(x= `PROX_CBD`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_CHILDCARE <- ggplot(data=condo_resale.sf, aes(x= `PROX_CHILDCARE`)) + 
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_ELDERLYCARE <- ggplot(data=condo_resale.sf, aes(x= `PROX_ELDERLYCARE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_URA_GROWTH_AREA <- ggplot(data=condo_resale.sf, 
                               aes(x= `PROX_URA_GROWTH_AREA`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_HAWKER_MARKET <- ggplot(data=condo_resale.sf, aes(x= `PROX_HAWKER_MARKET`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_KINDERGARTEN <- ggplot(data=condo_resale.sf, aes(x= `PROX_KINDERGARTEN`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_MRT <- ggplot(data=condo_resale.sf, aes(x= `PROX_MRT`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_PARK <- ggplot(data=condo_resale.sf, aes(x= `PROX_PARK`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_PRIMARY_SCH <- ggplot(data=condo_resale.sf, aes(x= `PROX_PRIMARY_SCH`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_TOP_PRIMARY_SCH <- ggplot(data=condo_resale.sf, 
                               aes(x= `PROX_TOP_PRIMARY_SCH`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

ggarrange(AREA_SQM, AGE, PROX_CBD, PROX_CHILDCARE, PROX_ELDERLYCARE, 
          PROX_URA_GROWTH_AREA, PROX_HAWKER_MARKET, PROX_KINDERGARTEN, PROX_MRT,
          PROX_PARK, PROX_PRIMARY_SCH, PROX_TOP_PRIMARY_SCH,  
          ncol = 3, nrow = 4)
```

### Drawing Statistical Point Map

```{r}
tmap_mode("view")
tmap_options(check.and.fix = TRUE) 
tm_shape(mpsz.sf)+
  tm_polygons() +
tm_shape(condo_resale.sf) +  
  tm_dots(col = "SELLING_PRICE",
          alpha = 0.6,
          style="quantile") +
  tm_view(set.zoom.limits = c(11,14))
tmap_mode("plot")
```

# Hedonic Pricing Modelling in R

## Simple Linear Regression Method

```{r}
condo.slr <- lm(formula=SELLING_PRICE ~ AREA_SQM, data = condo_resale.sf)
```

```{r}
summary(condo.slr)
```

```{r}
ggplot(data=condo_resale.sf,  
       aes(x=`AREA_SQM`, y=`SELLING_PRICE`)) +
  geom_point() +
  geom_smooth(method = lm)
```

## Multiple Linear Regression Method

```{r}
corrplot(cor(condo_resale[, 5:23]), diag = FALSE, order = "AOE",
         tl.pos = "td", tl.cex = 0.5, method = "number", type = "upper")
```
