---
title: "06 (Part 2) ‘Local Measures of Spatial Autocorrelation’"
author: "Heng Kuan Xin"
date: 2024-09-21
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  freeze: true
  output: true
  warning: false
  error: false

toc: true
toc-depth: 4
---

# Overview:

(Copied from Exercise)

Local Measures of Spatial Autocorrelation (LMSA) focus on the relationships between each observation and its surroundings, rather than providing a single summary of these relationships across the map. In this sense, they are not summary statistics but scores that allow us to learn more about the spatial structure in our data. The general intuition behind the metrics however is similar to that of global ones. Some of them are even mathematically connected, where the global version can be decomposed into a collection of local ones. One such example are Local Indicators of Spatial Association (LISA). Beside LISA, Getis-Ord’s Gi-statistics will be introduce as an alternative LMSA statistics that present complementary information or allow us to obtain similar insights for geographically referenced data.

In this hands-on exercise, you will learn how to compute Local Measures of Spatial Autocorrelation (LMSA) by using **spdep** package. By the end to this hands-on exercise, you will be able to:

-   import geospatial data using appropriate function(s) of **sf** package,

-   import csv file using appropriate function of **readr** package,

-   perform relational join using appropriate join function of **dplyr** package,

-   compute Local Indicator of Spatial Association (LISA) statistics for detecting clusters and outliers by using appropriate functions **spdep** package;

-   compute Getis-Ord’s Gi-statistics for detecting hot spot or/and cold spot area by using appropriate functions of **spdep** package; and

-   to visualise the analysis output by using **tmap** package.

### The Study Area and Data

(Copied from Exercise)

Two data sets will be used in this hands-on exercise, they are:

-   Hunan province administrative boundary layer at county level. This is a geospatial data set in ESRI shapefile format.

-   `Hunan_2012.csv`: This csv file contains selected Hunan’s local development indicators in 2012.

# Import the necessary

## Import packages

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse)
```

### Import Data

```{r}
# geospatial polygon feature layer
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan") # Note that crs is in WGS84, need to convert (TO BE COMPLETED)
# statistics data
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")

```

```{r}
# do a relational join to add statistics to sf object
hunan <- left_join(hunan,hunan2012) %>%
  select(1:4, 7, 15)

```

### Visualising Regional Development Indicator

```{r}
equal <- tm_shape(hunan) +
  tm_fill("GDPPC",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Equal interval classification")

quantile <- tm_shape(hunan) +
  tm_fill("GDPPC",
          n = 5,
          style = "quantile") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Equal quantile classification")

tmap_arrange(equal, 
             quantile, 
             asp=1, 
             ncol=2)
```

# Local Indicators of Spatial Association(LISA)

Copied from Exercise:

Local Indicators of Spatial Association or LISA are statistics that evaluate the existence of clusters and/or outliers in the spatial arrangement of a given variable. For instance if we are studying distribution of GDP per capita of Hunan Provice, People Republic of China, local clusters in GDP per capita mean that there are counties that have higher or lower rates than is to be expected by chance alone; that is, the values occurring are above or below those of a random distribution in space.

In this section, you will learn how to apply appropriate Local Indicators for Spatial Association (LISA), especially local Moran’I to detect cluster and/or outlier from GDP per capita 2012 of Hunan Province, PRC.

## Computing Spatial Weights (Queens')

```{r}
wm_q <- poly2nb(hunan, 
                queen=TRUE)
summary(wm_q)
```

## Standardise Weights (by row)

```{r}
rswm_q <- nb2listw(wm_q, 
                   style="W", 
                   zero.policy = TRUE)
rswm_q
```

## Computing local Moran’s I

```{r}
fips <- order(hunan$County)
localMI <- localmoran(hunan$GDPPC, rswm_q)
head(localMI)
```

*localmoran()* function returns a matrix of values whose columns are:

-   Ii: the local Moran’s I statistics

-   E.Ii: the expectation of local moran statistic under the randomisation hypothesis

-   Var.Ii: the variance of local moran statistic under the randomisation hypothesis

-   Z.Ii:the standard deviate of local moran statistic

-   Pr(): the p-value of local moran statistic

```{r}
printCoefmat(data.frame(
  localMI[fips,], 
  row.names=hunan$County[fips]),
  check.names=FALSE)
```

## Mapping the local Moran’s I

```{r}
hunan.localMI <- cbind(hunan,localMI) %>%
  rename(Pr.Ii = Pr.z....E.Ii..)

tm_shape(hunan.localMI) +
  tm_fill(col = "Ii", 
          style = "pretty",
          palette = "RdBu",
          title = "local moran statistics",
          midpoint = NA) +
  tm_borders(alpha = 0.5)
```

## Mapping local Moran’s I p-values

```{r}
tm_shape(hunan.localMI) +
  tm_fill(col = "Pr.Ii", 
          breaks=c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
          palette="-Blues", 
          title = "local Moran's I p-values") +
  tm_borders(alpha = 0.5)
```

## Mapping both local Moran’s I values and p-values

```{r}
localMI.map <- tm_shape(hunan.localMI) +
  tm_fill(col = "Ii", 
          style = "pretty", 
          title = "local moran statistics") +
  tm_borders(alpha = 0.5)

pvalue.map <- tm_shape(hunan.localMI) +
  tm_fill(col = "Pr.Ii", 
          breaks=c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
          palette="-Blues", 
          title = "local Moran's I p-values") +
  tm_borders(alpha = 0.5)

tmap_arrange(localMI.map, pvalue.map, asp=1, ncol=2)
```

# Creating a LISA Cluster Map

## Scatter Plot

```{r}
nci <- moran.plot(hunan$GDPPC, rswm_q,
                  labels=as.character(hunan$County), 
                  xlab="GDPPC 2012", 
                  ylab="Spatially Lag GDPPC 2012")
```

## Plotting Moran scatterplot with standardised variable

```{r}
hunan$Z.GDPPC <- scale(hunan$GDPPC) %>% 
  as.vector 

nci2 <- moran.plot(hunan$Z.GDPPC, rswm_q,
                   labels=as.character(hunan$County),
                   xlab="z-GDPPC 2012", 
                   ylab="Spatially Lag z-GDPPC 2012")
```

## Preparing LISA map classes

```{r}

quadrant <- vector(mode="numeric",length=nrow(localMI))

# Next, derives the spatially lagged variable of interest (i.e. GDPPC) and centers the spatially lagged variable around its mean.
hunan$lag_GDPPC <- lag.listw(rswm_q, hunan$GDPPC)
DV <- hunan$lag_GDPPC - mean(hunan$lag_GDPPC) 

# This is follow by centering the local Moran’s around the mean.
LM_I <- localMI[,1] - mean(localMI[,1])

# Next, we will set a statistical significance level for the local Moran.
signif <- 0.05

# These four command lines define the low-low (1), low-high (2), high-low (3) and high-high (4) categories.
quadrant[DV <0 & LM_I>0] <- 1
quadrant[DV >0 & LM_I<0] <- 2
quadrant[DV <0 & LM_I<0] <- 3  
quadrant[DV >0 & LM_I>0] <- 4      

# Lastly, places non-significant Moran in the category 0.
quadrant[localMI[,5]>signif] <- 0
```

## Plotting LISA map

```{r}
hunan.localMI$quadrant <- quadrant
colors <- c("#ffffff", "#2c7bb6", "#abd9e9", "#fdae61", "#d7191c")
clusters <- c("insignificant", "low-low", "low-high", "high-low", "high-high")

tm_shape(hunan.localMI) +
  tm_fill(col = "quadrant", 
          style = "cat", 
          palette = colors[c(sort(unique(quadrant)))+1], 
          labels = clusters[c(sort(unique(quadrant)))+1],
          popup.vars = c("")) +
  tm_view(set.zoom.limits = c(11,17)) +
  tm_borders(alpha=0.5)
```

## Plotting together with the p-value plot

```{r}
gdppc <- qtm(hunan, "GDPPC")

hunan.localMI$quadrant <- quadrant
colors <- c("#ffffff", "#2c7bb6", "#abd9e9", "#fdae61", "#d7191c")
clusters <- c("insignificant", "low-low", "low-high", "high-low", "high-high")

LISAmap <- tm_shape(hunan.localMI) +
  tm_fill(col = "quadrant", 
          style = "cat", 
          palette = colors[c(sort(unique(quadrant)))+1], 
          labels = clusters[c(sort(unique(quadrant)))+1],
          popup.vars = c("")) +
  tm_view(set.zoom.limits = c(11,17)) +
  tm_borders(alpha=0.5)

tmap_arrange(gdppc, LISAmap, 
             asp=1, ncol=2)
```

# Hot Spot and Cold Spot Area Analysis

Copied from Exercise:

Beside detecting cluster and outliers, localised spatial statistics can be also used to detect hot spot and/or cold spot areas.

The term ‘hot spot’ has been used generically across disciplines to describe a region or value that is higher relative to its surroundings (Lepers et al 2005, Aben et al 2012, Isobe et al 2015).

### Getis and Ord’s G-Statistics

An alternative spatial statistics to detect spatial anomalies is the Getis and Ord’s G-statistics (Getis and Ord, 1972; Ord and Getis, 1995). It looks at neighbours within a defined proximity to identify where either high or low values clutser spatially. Here, statistically significant hot-spots are recognised as areas of high values where other areas within a neighbourhood range also share high values too.

The analysis consists of three steps:

-   Deriving spatial weight matrix

-   Computing Gi statistics

-   Mapping Gi statistics

## Deriving distance-based weight matrix

```{r}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])
latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])
coords <- cbind(longitude, latitude)
```

```{r}
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dists)
```

### Computing Fixed Distance Matrix

```{r}
wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
wm_d62
```

```{r}
wm62_lw <- nb2listw(wm_d62, style = 'B')
summary(wm62_lw)
```

### Computing adaptive distance weight matrix

```{r}
knn <- knn2nb(knearneigh(coords, k=8))
knn
```

```{r}
knn_lw <- nb2listw(knn, style = 'B')
summary(knn_lw)
```

# Computing Gi statistics

## Gi statistics using fixed distance

```{r}
fips <- order(hunan$County)
gi.fixed <- localG(hunan$GDPPC, wm62_lw)
# gi.fixed
hunan.gi <- cbind(hunan, as.matrix(gi.fixed)) %>%
  rename(gstat_fixed = as.matrix.gi.fixed.)
```

## Mapping Gi values with fixed distance weights

```{r}
gdppc <- qtm(hunan, "GDPPC")

Gimap <-tm_shape(hunan.gi) +
  tm_fill(col = "gstat_fixed", 
          style = "pretty",
          palette="-RdBu",
          title = "local Gi") +
  tm_borders(alpha = 0.5)

tmap_arrange(gdppc, Gimap, asp=1, ncol=2)
```

## Gi statistics using adaptive distance

```{r}
fips <- order(hunan$County)
gi.adaptive <- localG(hunan$GDPPC, knn_lw)
hunan.gi <- cbind(hunan, as.matrix(gi.adaptive)) %>%
  rename(gstat_adaptive = as.matrix.gi.adaptive.)
```

## Mapping Gi values with adaptive distance weights

```{r}
gdppc<- qtm(hunan, "GDPPC")

Gimap <- tm_shape(hunan.gi) + 
  tm_fill(col = "gstat_adaptive", 
          style = "pretty", 
          palette="-RdBu", 
          title = "local Gi") + 
  tm_borders(alpha = 0.5)

tmap_arrange(gdppc, 
             Gimap, 
             asp=1, 
             ncol=2)
```
